A Method for the Spectroscopic Inference of Stellar Parameters

### 2014-06-14 tex/paper.tex

\documentclass[preprint]{aastex} %double-column, single-spaced document:
%\documentclass[iop,floatfix]{emulateapj}

\usepackage[breaklinks,colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}
\usepackage{apjfonts}
\usepackage{enumerate}
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor} 
\usepackage[utf8]{inputenc}

@@

%\vt stands for ``vector theta''
\newcommand{\vt}{ {\bm \theta}}
\newcommand{\vg}{\vt_{\star, {\rm grid}}}
\newcommand{\vpp}{\vt_{\star, {\rm post}}}
\newcommand{\finst}{f_{\lambda, {\rm inst}}}
\newcommand{\fsynth}{f_{\lambda, {\rm synth}}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vtstar}{\vt_{\star}}
\newcommand{\vtcheb}{\vt_{\rm Cheb}}
\newcommand{\vtglobal}{\vt_{\rm global}}
\newcommand{\vtorder}[1]{\vt_{{\rm order}_{#1}}}
\newcommand{\vtorders}{\vt_{\rm orders}}
\newcommand{\vtline}[1]{\vt_{{\rm line}_{#1}}}
\newcommand{\vtlines}{\vt_{\rm lines}}
\newcommand{\vtcov}{\vt_{\rm cov}}
\newcommand{\fM}{ \vec{{\bm M}}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ \vec{{\bm D}}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}

@@

\title{A Method for the spectroscopic inference of fundamental stellar parameters}
\author{\today{}\\
\medskip
Ian~Czekala\altaffilmark{1} et al.
%Author2\altaffilmark{2},
}

\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street 
 MS 10, Cambridge, MA 02138}
%\altaffiltext{2}{Institution 2}
\email{iczekala@cfa.harvard.edu}

\section{Introduction}
\begin{itemize}
  \item Where our technique fits in the ecosystem of stellar methods.
  \item Applicable to any kind of spectrum (long-slit, echelle, infrared,
    flux-calibrated or not)
  \item Examples of fields where accurate and unbiased stellar parameters are
    crucial. Exoplanets. T Tauri stars.
\end{itemize}

\section{Method}

A single spectrum provides a wealth of information that can be used to infer the fundamental properties of a star: effective temperature $T_{\rm eff}$, surface gravity $\log g$, and metallicity $\Z$; together $\vg$. We wish to extract the maximal amount information from a given spectrum while also recognizing the limitations imposed by our stellar model and exposing any degeneracies between the parameters.  For example, if we model a large range of the spectrum at once (e.g., the full optical range at high resolution), then we must explicitly account for any covariances and biases introduced by improper calibration or inaccuracies in the model spectrum. We present a generative model which uses a synthetic stellar library as a backend and apply post-processing techniques to mimic the observed spectrum as a function of several stellar parameters and calibration parameters in order to determine the best fit. By forward modelling the data spectrum in a Bayesian framework, our modular approach can easily incorporate additional calibration artifacts and stellar complexity (e.g., veiling or Zeeman splitting) while still being true to the limitations of the dataset.  This data driven approach will enable us to learn how the underlying synthetic models may be improved and refined in an iterative manner.

The roadmap for our method is as follows.
 First, for a given set of fundamental stellar parameters $\vg$, we generate a synthetic
 spectrum by interpolating from a grid of synthetic spectra (\S\ref{subsec:synthetic}). 
Then, we post-process the interpolated spectrum for realism by using additional
 parameters to describe Doppler shift, projected rotational velocity,
 interstellar extinction, and residual flux calibration errors
 (\S\ref{subsec:postprocess}). 
We perform a pixel-by-pixel comparison of the generated spectrum with the
 data spectrum using a prescribed likelihood function
 (\S\ref{subsec:likelihood}) and a parametric treatment of the covariance
 between pixel residuals (\S\ref{subsec:covariance}). 
We use a Markov chain Monte Carlo (MCMC) Gibbs sampler to numerically explore
the posterior probability density and determine constraints on the stellar
and nuisance parameters (\S\ref{subsec:MCMC}). 
The benefit of our method is that it provides robust estimates of the stellar and nuisance
parameters with well-quantified uncertainties and parameter degeneracies. 
Our parameterization of the covariance between pixel residuals also yields a means to identify
 and quantify the key discrepancies between the data and model, and provides a path forward to 
 improve the stellar models in an iterative fashion (\S\ref{subsec:learning}).
 \todo{I really want to overload the word ``covariance'' and use it to describe both the covariance between $T_{\rm eff}$ and $\log g$ as well as the covariance structure of the pixel residuals themselves. However I think this might be confusing so I am using ``degeneracies'' for the former and reserving ``covariance'' for the later.}

\subsection{Synthetic Model Spectra}
\label{subsec:synthetic}
Many methods exist to generate a synthetic model of a stellar spectrum. Generally, when
a high quality spectrum is needed for a specific set of stellar parameters, one
runs a stellar atmosphere and radiative transfer code to synthesize the
spectrum. Unfortunately, it is computationally expensive to accurately generate
a spectrum over a wide range of wavelengths, and so these codes are generally
employed to generate libraries of stellar spectra, which exist over a range of
$\vg$ at specified points on a regular grid, typically covering the main sequence of stellar
types. Several high-quality libraries of synthetic spectra now exist
(\citealt{hwd+13}, Kurucz, and more are being developed for GAIA).

In a forward-modelling approach, which may make many calls to generate a
spectra of varying $\vg$, it is currently computationally impractical to
iterate the structure of the stellar atmosphere and then run the stellar
synthesis for each call of $\vg$.  Some methods compromise by taking a halfway
approach, whereby a precomputed library of stellar atmospheres is interpolated
for a given $\vg$, and then a simple radiative transfer code is run to
synthesize a spectrum (e.g.  SME; \citealt{vp96}). This technique has the
advantage that interpolating atmospheres and then running the spectral
synthesis for each $\vg$ might be more accurate than directly interpolating
synthetic spectra. However, if a large region of wavelength is required at high
resolution, then the synthesis step will be slow.  A different approach might
be instead to eschew forward modelling (and thus repeated synthesis or grid
interpolation) of the spectra entirely and evaluate the posterior probability
density only at the grid points of the spectral library.  Then, these discrete
samples of the posterior density can be interpolated to find the posterior
surface and draw confidence intervals on the parameters (e.g. SPC;
\citealt{blj+12}). However, the difficulty with this approach is that the
parameter uncertainty might be smaller than the grid spacing, and then
interpolation will not properly capture degeneracies between parameters.

Because we desire a generative model of the spectra without the added
complexity or computational overhead of running a stellar synthesis, we opt to
use a \todo{Implement a more accurate but still fast interpolator for the grid.
Likely this will be a combination of using splines to pre-interpolate a finer
grid (~20 K, 0.1 dex spacing in gravity and metallicity and then doing linear
interpolation off of this new grid during the actual MCMC run. It will be described here.} However, the
techniques we develop in this paper to account for and parameterize the
correlated, systematic pixel residuals are generally applicable to spectra fit
by any of these methods.

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[width=0.85\textwidth]{figs/model_data.pdf}
  \caption{\textbf{Top}: Chebyshev polynomial which multiplies the model spectrum
   to account for inaccuracies in the flux-calibration. 
  \textbf{Middle}: The data spectrum and model spectrum, after it has been
   interpolated, post-processed, and multiplied by the Chebyshev polynomial.
  \textbf{Bottom}: Residuals from the model fit. Note the large residual at
   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{2}.}
\label{fig:model_data}
\end{center}
\end{figure*}

\subsection{Post-Processing}
\label{subsec:postprocess}

Libraries of synthetic spectra are generally computed at high resolution ($R
 \geq $100,000), sampled with many pixels per resolution element, and do not
 include any rotational broadening or account for any instrumental effects. 
In order to transform a raw synthetic spectrum interpolated from the grid into
 one that matches the spectrum of a real star, we must post-process the spectrum
 to account for these secondary effects. 
In addition to these fundamental parameters, a star also has several observed
 properties that are a function of its kinematics, geometry, and location in our
 galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$,
 total extinction $A_V$, and solid angle $\Omega$. 
We call these secondary parameters $\vpp$, because we can model their effects
 on the synthetic spectra in a ``post-processing'' step by convolution with an
 appropriate kernel or multiplication with a smooth function. 
Together, we call $\vtstar = \{\vg, \vpp\}$.

The projected rotational velocity of the star, parameterized by $v \sin i$,
broadens spectral line profiles and is mathematically described by convolution
with a specific kernel \citep[Equation 18.14]{gra08}. UV, optical, and infrared
spectra are acquired using a spectrograph which disperses light onto a CCD,
defining a specific resolution and sampling rate for the spectrum.  Because the
resolution of the spectrograph and pixel sampling are different from the
synthetic spectral library, we must convolve the raw spectrum with the line
spread function (LSF) of the instrument and resample to the exact pixels of the
CCD. In wavelength space, the $v \sin i$ and LSF operations are represented by
the convolution of the synthetic spectrum with these two kernels.  Using the
convolution theorem, we can rewrite these convolutions as multiplications in
Fourier space.  In order to use the fast Fourier transform (FFT) with discrete
samples of the synthetic spectrum, we must first resample it
onto a uniform grid.  In the case of a spectrum, the natural
uniform grid is one that is equally sampled in velocity space, such there is an
equal velocity shift $\Delta v$ between pixels. This results in a wavelength
grid that is linearly sampled in the logarithm of wavelength.  Next, we
multiply the Fourier transform of the synthetic spectrum by the Fourier
transforms of the rotational velocity and line spread function kernels.
Lastly, we do an inverse FFT to transform the modified spectrum back to
wavelength space, where it is sampled at the wavelength locations of the pixels
in the detector.  When resampling the spectrum, we are careful to ensure
band-limited interpolation using splines in order to prevent introducing
non-physical high frequency structure into the spectrum.

We scale the raw synthetic flux (often reported at the stellar surface) to an
observed flux by multiplying by the solid angle of the star, $\Omega =
R_\star^2/d^2$, where $R_\star$ is the stellar radius and $d$ is the distance
to the star from earth. Then, we correct for a radial velocity difference $v_z$ between the
star and the earth by Doppler shifting the model spectrum.  Finally, we correct
for interstellar extinction by dividing by a dust extinction model
$A_\lambda$, which is parameterized by extinction in the $V$ photometric band,
$A_V$. 
\todo{We can add in convolution with photometric filter
profiles to use photometry and provide a better constraint on the luminosity of
the star}
\todo{We might try implementing the convolution in wavelength space, rather
 than the FFT. 
This would allow the LSF to change with wavelength.\\}

\paragraph{Calibration polynomials} In contrast to a collection of photometric points in a spectral energy distribution, or any ensemble of data for which each data point is acquired by an individual measurement, adjacent wavelength points in a spectrum are measured simultaneously, with the same detector and--more importantly--the same calibration. Typically, spectra are flux calibrated by dividing by a ``sensitivity function,'' which is determined by comparing an observation of a spectrophotometric standard to a previously calibrated model spectrum. Therefore, any error in the sensitivity function, which is typically a smooth function with wavelength, will create a region of the data spectrum which is systematically offset by a small amount from what the true flux should be. The flux calibration error is then the residual difference in the empirically determined sensitivity function from the what the ``true'' sensitivity function would be if we had a perfectly calibrated instrument and spectrophotometric standard. Therefore, it is necessary to introduce additional degrees of freedom into our model in order to account for the calibration error and allow the model spectrum to properly match the data. By adding the residual calibration as a parameter in our model, we can explore the distribution of possible calibrations (\S\ref{subsec:MCMC}) and then marginalize the calibration parameters out of the final posterior distribution of the stellar parameters. This means that we can derive estimates of stellar parameters that properly account for the intrinsic uncertainty in the flux calibration.

A traditional way to deal with errors in the flux calibration is to avoid flux-calibration entirely and simply normalize both the data and the spectrum to the continuum level before comparison.  While this procedure may work well for hotter stars with well-defined continuua, normalizing spectra of cooler stars with large molecular features may introduce artificial features into the spectrum when the pseudo-continuum is incorrectly placed.  Instead, we choose to parameterize the residual flux calibration error by a low-order Chebyshev polynomial which is normalized to unity. The coefficients of the polynomial are parameters in the model that will be determined by the fitting algorithm, $\vtcheb = \{c_0, c_1, \ldots, c_N \}$. If we have multiple observations of spectrophotometric standards, then we can characterize the typical flux-calibration error of the spectrograph and put priors on the coefficients that will prevent overfitting of the polynomials.  For an example of the Chebyshev polynomial determined for a typical spectrum, see Figure~\ref{fig:model_data}. Using priors on the Chebyshev coefficients ensures that real spectral features, such as deep molecular bands in later type stars, are not accidentally treated as calibration artifacts. The overall normalization of the polynomial $c_0$ is degenerate with the solid angle of the star $\Omega$. Therefore we apply an additional constraint and enforce that the mean of the Chebyshev polynomial be equal to $1$. If there is only one order in the spectrum, such as with a long-slit spectrum or a single order of an echelle spectrum, then this means setting $c_0 = 1$. If we are fitting more than one order of an echelle spectrum then we adopt an independent Chebyshev polynomial for each order. This means that we chose the $c_0$ of a specific order to anchor to 1, while the $c_0$ for other orders are allowed to take on perturbations about $1$.

This polynomial formalism can in principle be used to model raw (not flux calibrated) spectra by using the Chebyshev polynomial to solve directly for the sensitivity function itself, rather than a residual calibration error as is currently done. When used in this manner, there will be greater uncertainty about the properties of the spectrograph, which will translate into wider priors on the Chebyshev coefficients.  In this case, there is more of a chance that the polynomial might destroy broad-scale information present in the spectrum by fitting it out, something that would be avoided with more accurate priors provided by a flux calibration.

\subsection{Likelihood Calculation}
\label{subsec:likelihood}

To evaluate which parameters of the model spectrum fit the data spectrum best,
we use a likelihood function to compare individual pixels between the two
spectra. A likelihood function provides the probability of the data spectrum
$\fD$, given the model spectrum $\fM(\vt)$, which is parameterized by the
stellar and calibration parameters $\vt$. The data spectrum and model spectrum are both composed of $N$ pixels, and so $\fD$ and $\fM$ represent the collection of these pixels in vector form, 
$\fD = \{\fDi\}^N_{i=1}$ and $\fM = \{\fMi\}^N_{i=1} $, respectively. The vector of residuals is the difference between the data spectrum and the model spectrum
\begin{equation}
  \fR(\vt) = \fD - \fM(\vt)
\end{equation}
In the language of probability calculus, the likelihood function is written as
$p(\fD | \fM(\vt))$, or $p(\fD | \vt)$ for short. We choose a standard
multidimensional Gaussian likelihood function which allows for covariance
between the pixel residuals 
\begin{equation}
  p(\fD | \vt) = \frac{1}{\sqrt{(2 \pi)^N \det(C)}} \exp\left ( -\frac{1}{2}
   \fR^T C^{-1} \fR \right ) 
   \label{eqn:prob}
\end{equation}
which penalizes models that yield larger residuals. $C$ is the covariance
matrix which describes the covariance between the pixel residuals. The structure of the covariance matrix is discussed in detail in the next section (\S\ref{subsec:covariance}). Because
the likelihood function generally exhibits a large dynamic range, it is numerically
preferred to use the log-likelihood function
\begin{equation}
  \ln[p(\fD | \vt)] = -\frac{1}{2} \fR^T C^{-1} \fR - \frac{1}{2} \ln \det C  
   - \frac{N}{2} \ln 2 \pi 
  \label{eqn:lnprob}
\end{equation}

\subsection{Parameterizing the covariance structure}
\label{subsec:covariance}
The covariance matrix $C$ describes the measurement noise in each pixel of the spectrum as well as the covariance between pixel residuals. It is fundamental to the likelihood calculation described in \S\ref{subsec:likelihood}. For most problems, such as fitting a spectral energy distribution to photometric measurements, each data point is acquired in an independent measurement from the other data points, which means that the noise for each data point, $\sigma_i$, is also independent from the other data points. This results in a diagonal covariance matrix
\begin{equation}
  C = 
  \begin{bmatrix}
    \sigma_0^2 & 0  & \cdots & 0\\
    0 & \sigma_1^2 & \cdots & 0\\
    \vdots  &   & \ddots  & \vdots \\
    0 & 0 & \cdots & \sigma_N^2\\
  \end{bmatrix}
  \label{eqn:covariance_diagonal}
\end{equation}
with $\sigma_{ij} = \delta_{ij} \sigma_{ij}$, where $\delta_{ij}$ is the Kronecker delta function. When the covariance matrix is diagonal, 
the likelihood function (Equation~\ref{eqn:lnprob}) reduces to the familiar
$\chi^2$ form of a sum over the square of the residuals, weighted by the
inverse variance of each data point
\begin{equation}
  \ln[p(\fD | \vt)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i^N
   \frac{R_i^2}{\sigma_i^2}
   \label{eqn:chi2}
\end{equation}

>>>>

\documentclass[preprint]{aastex} 
%\documentclass[iop,floatfix]{emulateapj}

\usepackage[breaklinks,colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}
%\usepackage{apjfonts}
\usepackage{enumerate}
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor} 
\usepackage[utf8]{inputenc}

@@

%\vt stands for ``vector theta''
\newcommand{\vt}{ {\bm \theta}}
\newcommand{\vT}{ {\bm \Theta}}
\newcommand{\vg}{\vt_{\star, {\rm grid}}}
\newcommand{\vpp}{\vt_{\star, {\rm post}}}
\newcommand{\finst}{f_{\lambda, {\rm inst}}}
\newcommand{\fsynth}{f_{\lambda, {\rm synth}}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vtstar}{\vt_{\star}}
\newcommand{\vtcheb}{\vt_{\rm Cheb}}
\newcommand{\vtglobal}{\vt_{\rm global}}
\newcommand{\vtorder}[1]{\vt_{{\rm order}_{#1}}}
\newcommand{\vtorders}{\vt_{\rm orders}}
\newcommand{\vtline}[1]{\vt_{{\rm line}_{#1}}}
\newcommand{\vtlines}{\vt_{\rm lines}}
\newcommand{\vtcov}{\vt_{\rm cov}}
\newcommand{\vM}{\mathsf{M}}
\newcommand{\vD}{\mathsf{D}}
\newcommand{\vR}{\mathsf{R}}
\newcommand{\vC}{\mathsf{C}}
\newcommand{\fM}{ \vec{{\bm M}}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ \vec{{\bm D}}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}

@@

\slugcomment{draft: \today{}}

\title{A Method for the Spectroscopic Inference of Stellar Parameters}
\author{Ian Czekala, Sean M. Andrews, et al.}
\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
%\email{iczekala@cfa.harvard.edu}

\section{Introduction \label{sec:intro}}

\comm{The following are some suggested paragraph outlines for a concise introduction:}

\comm{(1) start general: vast amount of information in an astronomical spectrum, and we want to 
make the most of that to learn about various physical parameters of interest.  accurate, unbiased 
stellar parameters (like temperature, gravity, metallicity, etc.) are of crucial importance in many 
astrophysics problems; e.g., exoplanets (and explain why), etc. (pick another one or two).}

\comm{(2) point out the challenge: quantitative extraction of these physical parameters is not 
easy.  go through some of the basic underlying limitations (practical, computational, limited 
models, etc.)}

\comm{(3) highlight various techniques/approaches being used (lots of references here).}

\comm{(4) now, without criticizing those approaches, just launch into a brief summary of our goals 
and a basic outline of the paper.}

\comm{I have made a lot of edits in Section 2, but have tried to restrict comments.  When there is 
something useful to say, I have added some commented-out text in the .tex file.}

\comm{Important: Note how I have changed the naming convention in the .bib file, so that a paper is 
referenced by the first author's last name and the year, like czekala14.  Your haphazard formatting 
was driving me nuts, and you can't find anything in the bib file (plus, if you read inside the .tex 
file, my approach makes things a hell of a lot easier).}

\section{Methodology \label{sec:method}}

All astronomers recognize that spectroscopy offers a wealth of information that can help 
characterize the fundamental properties of the observing target.  However, as noted in Section 
\ref{sec:intro}, the reliable and statistically robust inference of those physical (or empirical) 
parameters from an observed spectrum can be extraordinarily challenging.  Here we describe a 
generative Bayesian modeling framework that confronts some of the key obstacles in that process.  
The goal of this approach is to conservatively extract the maximal amount of information about a 
prescribed (and degenerate) parameter set by forward-modeling an observed spectrum, while also 
recognizing and explicitly accounting for the covariances and biases introduced by pathologically 
imperfect models or calibrations.  The method is modular, and therefore can easily incorporate 
additional physical or nuisance parameters as desired without sacrificing an accurate reflection of 
the limitations in the data.  Moreover, with a well-crafted observational sample, this data-driven 
approach should ultimately enable us to systematically learn how synthetic spectral models can be 
improved.  The specific applications discussed here are related to the spectra of individual stars, 
but the methodology is generic (and could be used for composite spectra of unresolved stellar 
clusters, galaxies, etc.).

The remainder of this section describes the mechanics of this modeling framework.  First, a model 
spectrum is generated for a given set of physical parameters (Section \ref{subsec:synthetic}), and 
then post-processed to mimic reality using a set of observational and practical nuisance parameters 
(Section \ref{subsec:postprocess}).  Next, a direct, pixel-by-pixel comparison between the data and 
model spectra is made with a prescribed likelihood function and a parametric treatment of the 
covariances between pixel residuals (Section \ref{subsec:likelihood}).  That process is iterated 
under the guise of hierarchical Monte Carlo Markov Chain (MCMC) simulations to numerically explore 
the posterior probability density of the model conditioned on the data, and thereby to determine 
constraints on the parameters of interest (Section \ref{subsec:MCMC}).  Along the way, these 
procedures are illustrated with real observations of the high resolution optical spectrum from a 
nearby F star.  That specific application, along with some alternative demonstrations of the 
method, are discussed in more detail in Section \ref{sec:examples}.

\subsection{Generating a Model Spectrum \label{subsec:synthetic}}

There are various approaches to synthesizing a spectrum, $f_{\lambda}$, for a specific set of model
parameters, $\vt_{\ast}$.  In an ideal case, a model stellar atmosphere is constructed and then 
subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
However, in general this approach is still computationally prohibitive for any iterative method of 
probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
structures that were pre-computed for a discrete grid of parameter values, $\{\vt_{\ast}\}^{\rm 
grid}$, for some arbitrary $\vt_{\ast}$, and then perform a radiative transfer calculation with 
that interpolated atmosphere to synthesize $f_{\lambda}$ \citep[e.g., as for SME;][]{vp96}.  A more 
common variant is to instead rely on interpolation over a library of pre-synthesized model spectra, 
$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While 
technically the former approach is most similar to the ideal case, the computational cost of 
repeated spectral synthesis is sufficiently high to make a detailed exploration of parameter space 
(particularly for data with a large spectral range) considerably less appealing.  A related, but 
different approach is to eschew forward modeling entirely (and therefore repeated spectral 
syntheses and/or library interpolations), and instead evaluate the models only at the discrete grid 
points of the library.  Then, these discretized samples of the posterior probability density can be 
interpolated to an arbitrary $\vt_{\ast}$ to construct appropriate confidence intervals 
\citep[e.g., the method of SPC;][]{blj+12}.  The difficulty with this latter approach is that the 
parameter uncertainties can be smaller than the grid spacing; in that case, there is valid concern 
that this interpolation might not accurately recover intrinsic parameter degeneracies.

Here we opt to take the computationally expedient approach that employs a library of model spectra, 
$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
(in practice, the metallicity $Z$ is often parameterized by the iron abundance, [Fe/H]).  However, 
it is worth noting that the techniques we will develop are applicable to {\it any} approach for 
generating a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
$\vt_{\ast}$ must be interpolated from among the spectral library, 
\begin{equation}
f_{\lambda}(\{\vt_{\ast}\}^{\rm grid}) \leadsto f_{\lambda}(\vt_{\ast}),
\label{eqn:interp}
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
with the corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).

\subsection{Post-Processing \label{subsec:postprocess}}

Generally, the ``raw" model spectrum $f_{\lambda}(\vt_{\ast})$ will be highly over-sampled compared 
to a typical observed spectrum, and does not account for several additional observational and 
instrumental effects that become important in comparisons with real data.  Therefore, a certain 
amount of post-processing is required before assessing the model quality.  We treat that 
post-processing in two stages: the first deals with an additional set of ``observational" 
parameters, $\vt_{\rm obs}$, that incorporate dynamical effects, geometry, and the relative 
location of the target, while the second employs a suite of nuisance (hyper-)parameters, $\vT_{\rm 
n}$, designed to mitigate an imperfect data calibration.

We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
former, we consider three kernels that contribute to the observed line-of-sight velocity 
distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, treats the instrumental 
spectral broadening.  For illustrative purposes we assume $\mathcal{F}_v^{\rm inst}$ is a Gaussian 
with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
a convolution in velocity-space,
\begin{equation}
f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) = f_{\lambda}(\vt_{\ast}) \otimes \varphi_v = f_{\lambda}(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
\label{eqn:broadening}
\end{equation}
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation}
f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) \mapsto \vM(\vt_{\ast}, \vt_{{\rm obs}, v}),
\label{eqn:resampling}
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-dimensional array $\vM$ (and $N_{\rm pix}$ is the number of pixels in the spectrum).  
For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
post-processing steps.
% SA: My idea for this figure is sort of 1-column wide (~3.5"), with a set of spectral panels going 
% down vertically.  The top panel shows a segment of the full resolution model.  The next one down 
% can show the individual kernels described above, as well as their composite, \varphi_v.  Then the 
% next panel down can show the convolved (but still full-resolution) spectrum.  Finally, the bottom 
% panel will show the spectrum re-sampled onto the observed pixels.  It need not be a large figure 
% (maybe 5" long?).

At this stage, the model is further modified in the flux dimension.  A typical synthetic spectrum 
is computed as the flux that would be measured {\it at the stellar surface}, and so needs to be 
diluted by the subtended solid angle, $\Omega = (R_{\ast}/d)^2$, where $R_{\ast}$ is the stellar 
radius and $d$ is the distance.  An additional wavelength-dependent scaling factor is applied to 
account for interstellar extinction, assuming a 
derived extinction law $A_{\lambda}$ \citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  
The parameters $\vt_{{\rm obs}, f} = [\Omega, \,\, A_V]$ are applied as
\begin{equation}
\vM(\vt) \equiv \vM(\vt_{\ast}, \vt_{\rm obs}) = \vM(\vt_{\ast}, \vt_{{\rm obs}, v}) \times \Omega \times 10^{-0.4\,A_{\lambda}},
\label{eqn:scaling}
\end{equation}
where we have simplified the notation by writing $\vt \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$.

So far, the procedure summarized in Eq.~\ref{eqn:interp}-\ref{eqn:scaling} is composed of 
relatively straightforward operations demanded by practical astronomical and computing issues.  If 
the observed spectrum was {\it perfectly} calibrated, we could proceed to a likelihood calculation 
(Section~\ref{subsec:likelihood}) that makes a direct comparison with $\vM(\vt)$ at this point.  
However, that is unlikely to be the case.  The primary concern is that an imperfect calibration 
produces (presumably low-level) mismatches in the underlying shape of the observed spectrum on 
relatively broad wavelength scales.  When compared with the model, these mismatches might represent 
a non-trivial contribution to the likelihood, and thereby bias our estimates of the desired 
physical parameters.  Often, this concern is treated externally to any modeling procedure, usually 
by dividing the observed spectrum (and model spectrum) by an appropriate polynomial function.  But 
that ``normalization" procedure implicitly assumes that there is no relevant information content on 
those mismatched scales; if $\vt$ also contributes to the broad spectral shape, then adopting this 
approach will corrupt the inferences of these parameters.  Moreover, in practice this approach is 
limited, since defining an appropriate polynomial becomes difficult in cases where the spectral 
line density is high (e.g., molecular bands for cool stars).

We adopt a somewhat analogous approach to deal with this issue, but cast it internal to the 
modeling framework to appropriately propagate the uncertainty introduced by additional degrees of 
freedom in the model.  The residual calibration errors are treated as an explicit contribution to 
the model spectrum, enabling us to explore the distribution of possible calibrations (see 
Section~\ref{subsec:MCMC}) and then eventually marginalize out the associated nuisance parameters.  
In essence, the inferences for the relevant stellar parameters will properly account for the 
underlying uncertainty in the calibration process.  This is achieved in practice by distorting the 
model spectrum with a (low-order) Chebyshev polynomial \citep[e.g.,][]{elh+06,koleva09}, 
\begin{equation}
\vM(\vt, \vT_{\rm cheb}) = \vM(\vt) \times \sum_n c_n \, T_n,
\end{equation}
where $T_n$ are the standard Chebyshev functions of order $n$ and the coefficients are treated as a
set of nuisance (hyper-)parameters, $\vT_{\rm cheb} = \{c_n\}^{\rm order}$, for each available 
spectral order in the dataset.  With judicious priors on $\vT_{\rm cheb}$, we can ensure that the 
unintended treatment of real spectral features (e.g., broad, deep molecular bands) as calibration 
artifacts is negligible (see Secton~\ref{sec:examples} for examples).  The lowest-order (scaling) 
coefficient, $c_0$, is by its nature degenerate with the solid angle parameter, $\Omega$.  
Therefore, we enforce an additional constraint by requiring that the mean of the polynomial is 
unity.  For data with a single spectral order, this means setting $c_0 = 1$.  If the goal is to 
model multiple spectral orders, we assign $c_0 = 1$ in an arbitrary order as an anchor, but permit 
the $c_0$ values in other orders to be different (as necessary).  It is worth noting that this 
formalism can, in principle, be extended to develop models for completely uncalibrated spectra, 
rather than the residual calibration mismatches as described here (with a suitable relaxation of 
the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:cheb} offers a practical demonstration of how 
these nuisance parameters are applied. 
% SA: I envision this figure much like the previous one, in a 1-column vertical distribution of 
% panels.  At top, you would show the data spectrum at an overlap of two spectral orders, along 
% with the post-processed model with no Chebyshev polynomials applied.  Next, you would show the 
% two Chebyshev polynomials appropriate for the two orders.  And then in the final panel you would 
% revisit the top panel with the polynomials applied, demonstrating the improvement.

\subsection{Model Evaluation \label{subsec:likelihood}}
